---
title: "parallel"
author: "Brian Mahardja"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Parallelizing training

Running bootstrapping and logistic regression in a serialized manner

```{r}
x <- iris[which(iris[,5] != "setosa"), c(1,5)]
trials <- 10000
res <- data.frame()
system.time({
  trial <- 1
  while(trial <= trials) {
    ind <- sample(100, 100, replace=TRUE)
    result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
    r <- coefficients(result1)
    res <- rbind(res, r)
    trial <- trial + 1
  }
})
```
Using lapply function
```{r}
x <- iris[which(iris[,5] != "setosa"), c(1,5)]
trials <- seq(1, 10000)
boot_fx <- function(trial) {
  ind <- sample(100, 100, replace=TRUE)
  result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
  r <- coefficients(result1)
  res <- rbind(data.frame(), r)
}
system.time({
  results <- lapply(trials, boot_fx)
})
```

Using lapply and parallelizing the function (running things in parallel)

```{r}
library(parallel)
library(MASS)
```

Figure out number of cores on our machine
```{r}
starts <- rep(100, 40)
fx <- function(nstart) kmeans(Boston, 4, nstart=nstart)
numCores <- detectCores()
numCores
```
Using regular lapply

User: how much the CPU takes to process 
System: how much time it takes for the background process
Elapsed: How long it takes to run the full function
```{r}
system.time(
  results <- lapply(starts, fx)
)
```


Use 'mclapply' to run on multiple cores
```{r}
system.time(
  results <- mclapply(starts, boot_fx, mc.cores=numCores)
)
```
Now let’s demonstrate with our bootstrap example:


```{r}
x <- iris[which(iris[,5] != "setosa"), c(1,5)]
trials <- seq(1, 10000)
boot_fx <- function(trial) {
  ind <- sample(100, 100, replace=TRUE)
  result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
  r <- coefficients(result1)
  res <- rbind(data.frame(), r)
}
system.time({
  results <- mclapply(trials, boot_fx, mc.cores = numCores)
})
```


```{r}
library(foreach)

for (i in 1:3){print(sqrt(i))
}

foreach (i=1:3) %do% {
  sqrt(i)
}
```
For parallel doeach

```{r}
library(foreach)
library(doParallel)

foreach (i=1:3) %dopar% {
  sqrt(i)
}

foreach (i=1:3, .combine=rbind) %dopar% {
  sqrt(i)
}
```

